---
title: "Report for Capstone Project about the Bank Marketing Dataset in XAI"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
author : Tomy Le
date : July 9, 2025
---

# 1. Exploratory Data Analysis (EDA)

## 1.1. Setup

First, let's import the necessary libraries for data manipulation and visualization. The `tidyverse` package includes `dplyr` (for data manipulation), `readr` (for reading data), `ggplot2` (for plotting), and `forcats` (for working with factors). We also load `gridExtra` to arrange plots.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(gridExtra) # For arranging multiple plots
library(stringr)   # For string manipulation
```

## 1.2. Initial Data Inspection

Let's get a first look at the data's structure, content, and statistical summary. For that, define the paths.

```{r}
# Define file paths
bank_path <- "data/bank/bank.csv"
bank_full_path <- "data/bank/bank-full.csv"
bank_additional_path <- "data/bank-additional/bank-additional.csv"
bank_additional_full_path <- "data/bank-additional/bank-additional-full.csv"

# Load the datasets
# read_delim is used for semicolon-separated files, similar to pd.read_csv(sep=';')
bank_df <- read_delim(bank_path, delim = ";", show_col_types = FALSE)
bank_full_df <- read_delim(bank_full_path, delim = ";", show_col_types = FALSE)
bank_additional_df <- read_delim(bank_additional_path, delim = ";", show_col_types = FALSE)
bank_additional_full_df <- read_delim(bank_additional_full_path, delim = ";", show_col_types = FALSE)
```

Let's inspect the dataframes.

```{r}
bank_df
bank_full_df
bank_additional_df
bank_additional_full_df
```
**Verification Conclusion:** The shapes match the documentation, and the smaller files are confirmed to be true subsets of the larger files. We can now confidently proceed with `bank-additional-full.csv` for our analysis.

## 1.3. Data Cleaning

We'll check for any missing values and duplicates. For that we will need to see what column entries exists for each column in the first place.
```{r}
# Loop through each column to check for unique values
for (col_name in colnames(bank_additional_full_df)) {
  
  unique_vals <- unique(bank_additional_full_df[[col_name]])
  num_unique_vals <- length(unique_vals)
  
  cat(paste0("\n----- Column: '", col_name, "' -----\n"))
  cat(paste0("Number of unique values: ", num_unique_vals, "\n\n"))
  
  # Set a threshold to decide whether to print all unique values
  if (num_unique_vals < 15) {
    cat(paste0("Unique values: ", paste(sort(unique_vals), collapse = ", "), "\n"))
  } else {
    cat(paste0("Values: [High Cardinality Feature - Sample: ", paste(head(unique_vals, 5), collapse = ", "), "...]\n"))
  }
}
```
## 1.4. Feature Analysis: Feature vs. Target

Now we'll analyze how each feature relates to the subscription outcome `y`. This will help us identify potentially predictive features.

### A. Categorical Features vs. Target ('y')
```{r}
# List of key categorical features to analyze
cat_features_to_plot <- c('job', 'marital', 'education', 'default', 'housing', 'loan', 'contact', 'poutcome')

# Create a list to store the plots
plot_list <- list()

for (col in cat_features_to_plot) {
  p <- ggplot(bank_additional_full_df, aes(y = fct_rev(fct_infreq(.data[[col]])), fill = y)) +
    geom_bar(position = "stack") +
    labs(title = paste0('"', str_to_title(col), '" vs. Subscription'), y = "", x = "Count", fill = "Subscribed") +
    theme_minimal() +
    scale_fill_viridis_d()
  
  plot_list[[col]] <- p
}

# Arrange and display plots using patchwork (part of tidyverse)
#patchwork::wrap_plots(plot_list, ncol = 2)
```


